---
title: "3D Plots for Option Greeks"
subtitle: "Visualising the change in option dynamics over time, strike and spot"
output:
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---
# **Setup Required Libraries**

```{r message=FALSE,warning=FALSE}
library(plotly)
```

# **Option Price and Implied Volatility**

We must set up functions for the price and implied volatility as parameters used to calculate and generate outputs for our 3D plots.


```{r}
BSprice<-function(pc, S, k, vol, d, r, t)
{

  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  
  BSprice = pc * exp(-d * t) * S * 
    pnorm(pc * d1) - pc * k * exp(-r * t) * pnorm(pc * d2)
  return(BSprice)
}


BSvol<-function(pc, S, k, price, d, r, t, start = 0.2)
{
  
  voli = start
  pricei = BSprice(pc, S, k, voli, d, r, t)
  vegai = BSvega(pc, S, k, voli, d, r, t)
  while(abs(price - pricei) > 0.000001) 
  {
    voli<-voli + (price - pricei) / vegai
    pricei<-BSprice(pc, S, k, voli, d, r, t)
    vegai<-BSvega(pc, S, k, voli, d, r, t)
  }
  
  BSvol = voli
  return(BSvol)
}
```

# **First-Order Greeks**

## Delta

Delta is the first derivative with respect to the underlying.

$$
\Delta = \frac{dV}{dS} = e^{-dt} \Phi(d_1)
$$

```{r}

BSdelta<-function(pc, S, k, vol, d, r, t)
{
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  
  if (pc == 1) {BSdelta = exp(-d * t) * pnorm(d1)} else 
  {BSdelta = exp(-d * t) * (pnorm(d1) - 1)}
  return(BSdelta)
}
```

## Gamma

The second derivative of the option value with respect to the underlying, Gamma, measures convexity in option value.

$$
\Gamma = \frac{d\Delta}{dS} = e^{-dt} \Phi(d_1)
$$


```{r}
BSgamma<-function(pc, S, k, vol, d, r, t)
{
  
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  
  BSgamma = exp(-d * t) * exp((-d1 ^ 2) / 2) / (sqrt(2 * pi) * S * vol * sqrt(t))
  
  return(BSgamma)}
```

## Theta

Theta is the derivative with respect to time-to-maturity. 

$$
\theta = \frac{dV}{d\tau} = -re^{-d\tau}K\Phi(d_2)-\frac{S\sigma}{2\sqrt{\tau}}\phi(d_1)
$$
$$
\text{where } \phi(d_1) = \frac{1}{\sqrt{2\pi}}e^{-\frac{1}{2}d_1^2}
$$


```{r}
BStheta<-function(pc, S, k, vol, d, r, t) 
{
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  
  BStheta = -exp(-d * t) * exp((-d1 ^ 2) / 2) * S * vol / 
    (sqrt(2 * pi) * 2 * sqrt(t)) + pc * d * S * exp(-d * t) * pnorm(pc * d1) - pc * r * k * exp(-r * t) * pnorm(pc * d2)
  return(BStheta)
}
```

## Vega

Vega is the derivative with respect to volatility. 

$$
\nu = \frac{dV}{d\sigma} = S\phi(d_1)\sqrt{\tau}
$$

```{r}
BSvega<-function(pc, S, k, vol, d, r, t)
{
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  
  BSvega = exp(-d * t) * S * sqrt(t) * exp((-d1 ^ 2) / 2) / (sqrt(2 * pi))
  return(BSvega)
}

```

## Rho

Rho represents the derivative with respect to the risk-free interest rate.

$$
\rho = \frac{dV}{dr}\pm KTe^{-rT}\Phi(\pm d_2)
$$

```{r}
BSrho<-function(pc, S, k, vol, d, r, t)
{
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  
  BSrho = pc * k * t * exp(-r * t) * pnorm(pc * d2)
  return(BSrho)
}
```

# **Higher-Order Option Greeks**

## Charm

$$
Charm = -\frac{d^2V}{dSd\tau} = -\phi(d_1)\frac{2r\sqrt{\tau}-\sigma d_2}{2\sigma\tau}
$$

```{r}
BScharm <-function(pc, S, k, vol, d, r, t) 
{
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  
  if (pc == 1) {
    BScharm = -1*(-(1/(sqrt(2*pi)))*exp(-d1^2/2)*(2*r*sqrt(t)-vol*d2)/(2*vol*t))} else 
      {BScharm = -(1/(sqrt(2*pi)))*exp(-d1^2/2)*(2*r*sqrt(t)-vol*d2)/(2*vol*t)}
  
  return(BScharm)
}
```

## Vanna

$$
Vanna = \frac{d^2V}{d\sigma dS} = \sqrt{\tau} \phi(d_1) [1-d_1]
$$


```{r}
BSvanna<-function(pc, S, k, vol, d, r, t){ 
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSvanna2 = -BSgamma(pc, S, k, vol, d, r, t) * (sqrt(t) * S) * d2
  return(BSvanna2)
}
```


## VegaVanna

$$
VegaVanna = \frac{d^3V}{d\sigma^2 dS} 
$$

```{r}
BSvegavanna<-function(pc, S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSvegavanna2 = BSvanna(pc,S,k,vol,d,r,t) * (1 / vol) * (d1 * d2 - d1 / d2 - 1)
  return(BSvegavanna2)
}
```

## Veta

$$
Veta = \frac{d^2V}{d\sigma d\tau} 
$$

```{r}
BSveta<-function(S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSveta = -S * exp(-d * t) * exp((-d1 ^ 2) / 2) / (sqrt(2 * pi)) * sqrt(t) * (d + (((r-d) * d1) / (vol * sqrt(t))) - ((1 + d1 * d2) / 2*t))
  return(BSveta)
}
```

## Volga

$$
Volga = \frac{d^2V}{d\sigma^2 } 
$$

```{r}
BSvolga<-function(S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSvolga = exp(-d * t) * sqrt(t) * exp((-d1 ^ 2) / 2) / (sqrt(2 * pi)) * ((d1 * d2) / vol)
  return(BSvolga)
}
```

## Speed

$$
Speed = \frac{d^3V}{dS^3} 
$$

```{r}
BSspeed<-function(pc, S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSspeed = (-BSgamma(pc, S, k, vol, d, r, t) / S) * ((d1 / (vol * sqrt(t))) + 1)
  return(BSspeed)
}
```

## Zomma

$$
Zomma = \frac{d^3V}{d\sigma dS^2} 
$$

```{r}
BSzomma<-function(pc, S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSzomma = BSgamma(pc, S, k, vol, d, r, t) * ((d1 * d2 - 1) / vol)
  return(BSzomma)
}
```

## Ultima

$$
Ultima = \frac{d^2S}{d\sigma^3 } 
$$

```{r}
BSultima<-function(pc, S, k, vol, d, r, t){
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  BSultima = (-BSvega(pc, S, k, vol, d, r, t) / vol^2) * (((d1 * d2) * (1 - d1 * d2)) + d1^2 + d2^2)
  return(BSultima)
}
```

# **3D Plots**

## Parameters

We outline some base parameters for the options which we want to examine. We will start with simple ATM call options, and set up our x, y and z axes which are the time to maturity, strikes (or spot), and Greeks.

```{r}
pc = 1
S = 100;
T = 5;
r = .01;
vol = .15;
K = 100
div = 0

ttm = seq(0.1, T, .25)
strikes = seq(0,2*S,1)
s0 = seq(0,2*K,.5)
```

## Delta

We can generate plots as either a function of the strike:

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSdelta(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "delta"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

Or as a function of the spot:

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(S) BSdelta(pc,S,K,vol,div,r,ttm), s0))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "spot"),
                      zaxis = list(title = "delta"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

We can also save and store the outputted plots:

```{r}
# To save figure
# plt: the output from plot_ly()
# orca(plt, file="delta.eps")
```


## Theta

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BStheta(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                  scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "theta"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)
plt
```


## Gamma

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSgamma(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "gamma"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Vega

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSvega(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "vega"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Vanna

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSvanna(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "vanna"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Charm

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BScharm(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "charm"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Volga

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSvolga(S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "volga"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## VegaVanna

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSvegavanna(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "vegavanna"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Veta

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSveta(S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "veta"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Ultima

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSultima(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "ultima"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```

## Speed

```{r}
nrow <- length(ttm)
ncol <- length(strikes)
to_plot <- list(x = matrix(rep(ttm, ncol), nrow=nrow, ncol=ncol, byrow=FALSE),
                y = matrix(rep(strikes, nrow), nrow=nrow, ncol=ncol, byrow=TRUE),
                z = mapply(function(K) BSspeed(pc,S,K,vol,div,r,ttm), strikes))

plt<-plot_ly(x = to_plot$x, y = to_plot$y, z = to_plot$z) %>% 
  layout(margin = list(l=0, r=0, t=0, b=0), 
                 scene = list(xaxis = list(title = "time to maturity"),
                      yaxis = list(title = "strike"),
                      zaxis = list(title = "speed"),
                      camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)),
                      aspectratio = list(x = 1, y = 1, z = 1))) %>%
  add_surface(colorscale = list(seq(0,1,length.out=7),
                                c("#18389B","#02B1E8","#02B283","#FEBB08","#F93616","#F7103C","#CC0B8C")),
              showscale = FALSE)

plt
```


