---
title: "Option Pricing - Binomial and Trinomial Trees"
subtitle: Building trees and comparing calibration methods to the Black-Scholes model
author: "Gary He"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---


# Option tree models and calibration parameters

Whilst stock prices and hence option values can be viewed as variables that evolve in continuous time, it is possible to discretise these processes. Cox-Ross-Rubenstein (1979) and Jarrow-Rudd (1983) developed calibration parameters for pricing options using simple binomial trees.

By modelling option as simply going either 'up' or 'down' at each discrete step in time with probability p, we can approximate a continuous evolution of the price. The same can be applied to stocks, or any asset.

The CRR calibration parameters are:

$$
up = u = e^{\sigma \sqrt{h}} 
$$
$$
down = d = e^{-\sigma \sqrt{h}}
$$

$$
p = \frac{1}{2} + \frac{1}{2} (\frac{\mu}{\sigma}) \sqrt{h}
$$


The JR calibration parameters are:

$$
up = u = e^{\mu h + \sigma \sqrt{h}} 
$$

$$
down = d = u^{\mu h - \sigma \sqrt{h}}
$$

$$
p = \frac{1}{2}
$$


# Building binomial and trinomial trees


## Parameters

We can display the binomial and trinomial trees as matrices in R. First we set out some parameters:

```{r}
  Call_Put = 1;
  Eu_Am = 0;
  S = 30;
  K = 30;
  sigma = .4;
  delta = .05;
  r = .05;
  T_t = 5/12;
  
  Steps = c(15, 20, 25, 30, 50, 75, 100, 125, 150, 200, 250, 300, 500) 
  n = length(Steps)
```

Now we can begin plotting trees.

## Binomial tree

```{r}
optiontree <- function(u, d, p, n, discount, S, K, pc, ea)
{
  # ea=1 is American, ea=0 is European
  B <-matrix(0,ncol=n+1,nrow=n+1)

  # computing the final payoffs
  for (i in 1:(n+1))
  {B[i,n+1] <- max(pc * (S * (u^(n+1-i)) * (d^(i-1)) - K), 0)}
  
  # building the tree backwards
  for (i in 1:n)
    {
      for (j in (i+1):(n+1))
      {
      B[j, n+1-i] = (p * B[j-1, n-i+2] + (1-p) * B[j, n-i+2]) * discount;
  
      #for american options
      if (ea==1)
        {B[j, n+1-i] = max(B[j, n+1-i], pc * (S * (u^(n+1-j)) * (d^(j-i-1)) - K))}
  
      }
    }
  return(B)
}
```

## Trinomial tree

```{r}
Trinomial<-function(Spot, K, T, r, q, v, n, PutCall, EuroAmer)
{
  dt = T / n;
  u = exp(v * sqrt(2*dt));
  d = 1 / u;
  pu = (exp((r-q) * dt/2) - exp(-v * sqrt(dt/2)))^2 / (exp(v * sqrt(dt/2)) - exp(-v * sqrt(dt/2)))^2;
  pd = (exp(v * sqrt(dt/2)) - exp((r-q) * dt/2))^2 / (exp(v * sqrt(dt/2)) - exp(-v * sqrt(dt/2)))^2;
  pm = 1 - pu - pd;
  
  # Tree for stock price minus present value of dividends
  
  S <- matrix(0,nrow=2*n+1, ncol=n+1)
  S[1,1] = Spot;
  
  for (j in 2:(n+1))
  {
    for (i in 1:(2*j-1))
    {
    S[i,j] = S[1,1] * u^j * d^i;
    }
  }
  
  # Calculate Terminal Price for Calls and Puts
  Op <- matrix(0,nrow=2*n+1, ncol=n+1)
  for (i in 1:(2*n+1))
  {
    if(PutCall=='Call'){ Op[i, n+1] <- max(S[i, n+1] - K, 0)} else {
    if(PutCall=='Put'){ Op[i, n+1] <- max(K - S[i, n+1], 0)}}
  }
  
  # Calculate Remaining entries for Calls and Puts
  
  for (j in n:1)
  {
      for (i in 1:(2*j-1))
      {
      if (EuroAmer=='Amer'){
      if (PutCall == 'Call')
        {Op[i, j] = max(S[i, j] - K, exp(-r*dt) * (pu*Op[i, j+1] + pm*Op[i+1, j+1] + pd*Op[i+2, j+1]))} else 
          { if (PutCall == 'Put')
          {
                Op[i, j] = max(K - S[i, j], exp(-r*dt) * (pu*Op[i, j+1] + pm*Op[i+1, j+1] + pd*Op[i+2, j+1]))}
          }}
      if (EuroAmer=='Euro'){
        Op[i, j] = exp(-r*dt) * (pu*Op[i, j+1] + pm*Op[i+1, j+1] + pd*Op[i+2, j+1])}
      }
  }
  
  y = Op
  return(y)
       
}

Trinomial(S,K,T_t,r,delta,sigma,5,'Call','Euro')
```

## Calibrating and plotting the binomial trees - CRR

```{r}

CRRoptiontree<-function(pc, ea, S, K, vol, q, r, T, n)
{
  #computing the parameters of the binomial tree
  dt = T / n;
  u = exp(vol * sqrt(dt));
  d = 1 / u;
  p = (exp((r - q) * dt) - d) / (u - d);
  discount = exp(-r * dt);
  
  B = optiontree(u, d, p, n, discount, S, K, pc, ea);
  return(B)
}
  
    
CRRprice<-function(pc, ea, S, K, vol, q, r, T, n)
{
  B = CRRoptiontree(pc, ea, S, K, vol, q, r, T, n);
  y = B[n+1,1]; # The price is the first node, which in the tree is the bottom left element of the tree matrix
  return(y)
}

CRRoptiontree(Call_Put,Eu_Am,S,K,sigma,delta,r,T_t,5)
```

## Calibrating the binomial trees - JR


```{r}

JRoptiontree<-function(pc, ea, S, K, vol, q, r, T, n)
{
  #computing the parameters of the binomial tree
  dt = T / n;
  u = exp((r - q - (vol^2)/2) * dt + vol * sqrt(dt));
  d = exp((r - q - (vol^2)/2) * dt - vol * sqrt(dt));
  p = 0.5;
  discount = exp(-r * dt);
  B = optiontree(u, d, p, n, discount, S, K, pc, ea);
  return(B)
}

JRprice<-function(pc, ea, S, K, vol, q, r, T, n)
{
  B = JRoptiontree(pc, ea, S, K, vol, q, r, T, n);
  y = B[n+1,1]
  return(y)
}

JRoptiontree(Call_Put,Eu_Am,S,K,sigma,delta,r,T_t,5)
```

# Comparing convergence speeds of discrete trees against the Black-Scholes model

## Plotting the convergence

```{r}

BSprice<-function(pc, S, k, vol, d, r, t)
{
  
  
  d1 = (log(S / k) + t * (r - d + (vol ^ 2) / 2)) / (vol * sqrt(t))
  d2 = d1 - vol * sqrt(t)
  
  BSprice = pc * exp(-d * t) * S * 
    pnorm(pc * d1) - pc * k * exp(-r * t) * pnorm(pc * d2)
  return(BSprice)
}

CRR_Prices = rep(0,n);
for (i in 1:n) CRR_Prices[i] = CRRprice(Call_Put,Eu_Am,S,K,sigma,delta,r,T_t,Steps[i]);


JR_Prices = rep(0,n);
for (i in 1:n) JR_Prices[i] = JRprice(Call_Put,Eu_Am,S,K,sigma,delta,r,T_t,Steps[i]);


Tr_Prices = rep(0,n);
for (i in 1:n) Tr_Prices[i] = Trinomial(S,K,T_t,r,delta,sigma,Steps[i],'Call','Euro')[1,1];

BS_Prices = BSprice(Call_Put,S,K,sigma,delta,r,T_t) * rep(1,n);

plot(CRR_Prices,main='Convergence',xaxt="n",type="l",col="blue",xlab="Steps",ylab="Price")
axis(1,at=1:13,labels=Steps)
lines(JR_Prices,col="purple")
lines(Tr_Prices,col="green")
lines(BS_Prices,col="black")
legend("topright",c('CRR','JR','Trinomial','BS'),
       col=c("blue", "purple","green","black"),lty=c(1,1,1,1),cex=0.8);

```

